Computation and comparison of the stable northeastern United States marine boundary layer

### 2020-12-07 Paper/main.tex

\begin{abstract}
  In this work we investigate the behavior of stable marine boundary layers
  located near the coast of the Northeastern United States.  Using the ExaWind
  large eddy simulation (LES) codes, three stable atmospheric conditions were
  chosen to match the Cape Wind measurements of Archer \emph{et al.} with wind
  speeds of 5 m/s, 10 m/s, and 15 m/s at the 20 m measurement height.  The
  behavior of the stable boundary layers, including mean flow quantities and
  turbulent statistics, are examined and compared to previous computations of
  the neutral and unstable offshore boundary layer at the same location.  This
  study also examines the domain and mesh requirements necessary to capture the
  turbulent scales for the stable offshore boundary layers.  Finally, we compare
  solutions computed using both Nalu-Wind and AMR-Wind solvers, and report on
  their relative accuracy and performance in this study.

@@

Several recent measurement campaigns have provided data to understand
the ABL and wind characteristics for potential offshore wind farms in
the Northeastern United States.  Pichugina et al. \cite{pichugina2017properties}
measured the wind profiles and vertical shear profiles in the Gulf of
Maine using a ship-borne lidar approach.  Analysis of measured data
sets from Nantucket Sound by Archer et
al. \cite{archer2016predominance} showed a predominance of low-shear,
unstable conditions at the site.  However, strong seasonal variations
and stratification changes due to diurnal variation were also
observed.

@@

The first successful LES of the SBL was published by Mason and Derbyshire in
1990 \cite{Mason1990}. They used a strategy of simulating a near steady-state
neutral stability turbulent simulation, then added a cooling-surface boundary
to develop the stable ABL. They recorded a need for implausibly strong initial
perturbations to sustain turbulence using the Smagorinsky SGS model. Brown et
al.\ \cite{Brown1994} worked with Mason and Derbyshire to add stochastic backscatter
to the SGS model, resulting in better agreement with surface layer observations.
Andren \cite{Andren1995}, simulated two SBL types, one with a neutral layer aloft, and
one capped by an inversion, comparing two SGS models. He found significant improvements
with the Sullivan SGS model, on par with the earlier backscatter model. Kemp and
Thompson \cite{Kemp1996} used the same model as \cite{Brown1994} to study particle
dispersion, but the LES code did not replicate an expected plume growth. Derbyshire
wrote a review of SBL modeling along with future projections in 1999 \cite{Derbyshire1999},
noting the importance of local Richardson number and identifying a need to go
to higher resolutions in SBL research.

In 2000, Kosovic and Curry \cite{Kosovic2000} used an LES simulation
with a nonlinear SGS model containing backscatter to favorably compare
to the Beaufort Sea Arctic Stratus Experiment (BASE) measurements,
showing that the SBL could reach a quasi steady-state. They also
showed good agreement with the Nieuwstadt analytical model. In the same
year, Saiki et al. \cite{Saiki2000} simulated an SBL with
a strong capping inversion at z=500 m, producing reasonable surface layer profiles and
identifying a need for SBL specific SGS models. Beare and MacVean \cite{Beare2004}
presented the first grid-focused work on LES of the SBL, using grids with deltas
between 2 m and 12.5 m, finding significant variation between grid sizes. The inclusion
of a backscatter model resulted in significant variation at a 6.25 m resolution, but
very little at a 2 m resolution. Basu et al. \cite{Basu2005} published a set of
turbulence statistics, measurements, and LES simulations in support
of the viability of Nieuwstadt's local scaling hypothesis under very stable
conditions. Jimenez and Cuxart \cite{Jimenez2005} set out to quantify the range
for which standard SGS models, using the Kolmogorov theory for dissipation, were
valid for increasingly stable flows. They found the model to perform adequately
for weakly stable, and some moderately stable flow regimes.

An inter-comparison of LES simulations of the SBL was organized in 2006, the
results of which are published in Beare et al.\ \cite{Beare2006}. Collectively,
this is known as the GABLS initiative, and involved simulating a moderately
stable flow taken from Kosovic and Curry \cite{Kosovic2000}. Eleven codes
were compared in all, wherein grid deltas of less than 12.5 m were found to
successfully maintain resolved turbulence, grid deltas less than 3.125m
showed little statistical change, and results became independent of SGS
modeling at 1 m resolution. In the same year, Kumar et al. \cite{Kumar2006}
successfully simulated an entire diurnal cycle with LES using a Lagrangian
dynamic scale dependent SGS model. This model has been used by several
researchers since 2011, including Lu and Porte-Agel \cite{Lu2011}, Huang
and Bou-Zeid \cite{Huang2013}, Park et al.\ \cite{Park2014}, and
Abkar and Porte-Agel \cite{Abkar2015}. Stoll and Porte-Agel \cite{Stoll2007}
compared three grid refinements and three scale dependent dynamic SGS models,
showing that the Lagrangian and locally averaged models reproduced important
flow features. The work of van Dop and Axelsen \cite{vanDop2007} provided
even more evidence in support Nieuwstadt's local scaling hypothesis for SBLs.
Chow and Zhou \cite{Chow2011} showed evidence that the Smagorinsky model is
simply too dissipative for use in strongly stable flows.

@@

\subsection{Objectives}
In the current study on stable marine
boundary layers, we aim to build on the previous work in the
literature and complete several objectives.  First, we examine the
domain and mesh requirements necessary to capture the turbulent scales
for the stable offshore boundary layers.  Secondly, two LES solvers
within the ExaWind project will be compared in this study: the
unstructured, incompressible Navier-Stokes solver Nalu-Wind and the
structured, adaptive mesh refinement solver AMR-Wind.  Both codes are
used to calculate the same SBL, and the relative differences between
the codes are reported.  Finally, a series of stable boundary layer
conditions will be computed using AMR-Wind that match the
measurements of Archer et al. \cite{archer2016predominance}.  The
behavior of the stable boundary layers, including mean flow quantities
and turbulent statistics, are examined and compared to the neutral and
unstable cases computed in \cite{cheung2020large}.

>>>>

\begin{abstract}
  In this work we investigate the behavior of stable marine boundary layers
  located near the coast of the Northeastern United States.  Using the ExaWind
  large eddy simulation (LES) codes, three stable atmospheric conditions were
  chosen to match the Cape Wind measurements of Archer \emph{et al.} with wind
  speeds of 5 m/s, 10 m/s, and 15 m/s at the 20 m measurement height.  The
  behavior of the stable boundary layers, including mean flow quantities and
  turbulent statistics, are examined and compared to previous computations of
  the neutral and unstable offshore boundary layer at the same location.  This
  study also examines the domain and mesh requirements necessary to capture the
  turbulent scales for the stable offshore boundary layers.  Finally, we compare
  solutions computed using both Nalu-Wind and AMR-Wind solvers, and
  compare their predicted solutions and performance in this study.

@@

Several recent measurement campaigns have provided data to understand
the ABL and wind characteristics for potential offshore wind farms in
the Northeastern United States.  Pichugina et al. \cite{pichugina2017properties}
measured the wind profiles and vertical shear profiles in the Gulf of
Maine using a ship-borne LIDAR approach.  Analysis of measured data
sets from Nantucket Sound by Archer et
al. \cite{archer2016predominance} showed a predominance of low-shear,
unstable conditions at the site.  However, strong seasonal variations
and stratification changes due to diurnal variation were also
observed.

@@

The first successful LES of the SBL was published by Mason and Derbyshire in
1990 \cite{Mason1990}. They used a strategy of simulating a near steady-state
neutral stability turbulent simulation, then added a cooling-surface boundary
to develop the stable ABL. They recorded a need for implausibly strong initial
perturbations to sustain turbulence using the Smagorinsky SGS model. Brown et
al.\ \cite{Brown1994} worked with Mason and Derbyshire to add stochastic backscatter
to the SGS model, resulting in better agreement with surface layer observations.
Andren \cite{Andren1995} simulated two SBL types, one with a neutral layer aloft, and
one capped by an inversion, comparing two SGS models. He found significant improvements
with the Sullivan SGS model, on par with the earlier backscatter model. Kemp and
Thompson \cite{Kemp1996} used the same model as \cite{Brown1994} to study particle
dispersion, but the LES code did not replicate an expected plume growth. Derbyshire
wrote a review of SBL modeling along with future projections in 1999 \cite{Derbyshire1999},
noting the importance of local Richardson number and identifying a need to go
to higher resolutions in SBL research.

In 2000, Kosovic and Curry \cite{Kosovic2000} used an LES simulation
with a nonlinear SGS model containing backscatter to favorably compare
to the Beaufort Sea Arctic Stratus Experiment (BASE) measurements,
showing that the SBL could reach a quasi steady-state. They also
showed good agreement with the Nieuwstadt analytical model. In the same
year, Saiki et al. \cite{Saiki2000} simulated an SBL with
a strong capping inversion at $z$=500 m, producing reasonable surface layer profiles and
identifying a need for SBL specific SGS models. Beare and MacVean \cite{Beare2004}
presented the first grid-focused work on LES of the SBL, using grids with deltas
between 2 m and 12.5 m, finding significant variation between grid sizes. The inclusion
of a backscatter model resulted in significant variation at a 6.25 m resolution, but
very little at a 2 m resolution. Basu et al. \cite{Basu2005} published a set of
turbulence statistics, measurements, and LES simulations in support
of the viability of Nieuwstadt's local scaling hypothesis under very stable
conditions. Jimenez and Cuxart \cite{Jimenez2005} set out to quantify the range
for which standard SGS models, using the Kolmogorov theory for dissipation, were
valid for increasingly stable flows. They found the model to perform adequately
for weakly stable, and some moderately stable flow regimes.

An inter-comparison of LES simulations of the SBL was organized in 2006, the
results of which are published in Beare et al.\ \cite{Beare2006}. Collectively,
this is known as the GABLS initiative, and involved simulating a moderately
stable flow taken from Kosovic and Curry \cite{Kosovic2000}. Eleven codes
were compared in all, wherein grid deltas of less than 12.5 m were found to
successfully maintain resolved turbulence, grid deltas less than 3.125 m
showed little statistical change, and results became independent of SGS
modeling at 1 m resolution. In the same year, Kumar et al. \cite{Kumar2006}
successfully simulated an entire diurnal cycle with LES using a Lagrangian
dynamic scale dependent SGS model. This model has been used by several
researchers since 2011, including Lu and Porte-Agel \cite{Lu2011}, Huang
and Bou-Zeid \cite{Huang2013}, Park et al.\ \cite{Park2014}, and
Abkar and Porte-Agel \cite{Abkar2015}. Stoll and Porte-Agel \cite{Stoll2007}
compared three grid refinements and three scale dependent dynamic SGS models,
showing that the Lagrangian and locally averaged models reproduced important
flow features. The work of van Dop and Axelsen \cite{vanDop2007} provided
even more evidence in support Nieuwstadt's local scaling hypothesis for SBLs.
Chow and Zhou \cite{Chow2011} showed evidence that the Smagorinsky model is
simply too dissipative for use in strongly stable flows.

@@

\subsection{Objectives}
In the current study on stable marine boundary layers, we aim to build
on the previous work in the literature and complete several
objectives.  First, we examine the domain and mesh requirements
necessary to capture the turbulent scales for the stable offshore
boundary layers.  Secondly, two incompressible-flow LES solvers within
the ExaWind simulation suite will be compared in this study: The
unstructured-grid solver Nalu-Wind and the structured, adaptive mesh
refinement (AMR) solver AMR-Wind.  Both codes are used to calculate
the same SBL, and the relative differences between the codes are
reported.  Finally, a series of stable boundary layer conditions will
be computed using AMR-Wind that match the measurements of Archer et
al. \cite{archer2016predominance}.  The behavior of the stable
boundary layers, including mean flow quantities and turbulent
statistics, are examined and compared to the neutral and unstable
cases computed in \cite{cheung2020large}.

### 2020-12-07 Paper/results.tex
Minor change to sentence in results.tex

Additional information about the mesh and domain requirements can be
drawn by examining the turbulent integral length scale $L$.  This
length scale can be calculated from $\langle R_{11}(\boldsymbol{\xi})
\rangle$ via
\begin{equation}
  L = \int_0^\infty \langle R_{11}(\xi) \rangle \: {\textrm d}\xi,
\end{equation}
and the results are shown in Table \ref{tab:GridStudyLscale}.  With a
resolution of 2.5m, the ratio of the longitudinal length scale to the
grid size $L/\textrm{d}x \approx 10.1$, while in the lateral direction
$L/\textrm{d}x \approx 2.36$, suggesting even finer resolution is
required to capture the lateral scales.  This may be a critical
consideration when resolving both the lateral and longitudinal
turbulent scales are important, such as in SBL simulations with
turbine interactions.  However, in other situations, the additional
computational cost incurred may be prohibitive and only the
longitudinal component can be resolved.

@@

For both AMR-Wind and Nalu-Wind each time step is dominated by the linear solver work.
A time step within AMR-Wind requires two Poisson solves (MAC and nodal projection),
a tensor Helmholtz solve for Momentum equations, and two scalar Helmholtz solves
(potential temperature equation and subgrid-scale kinetic energy equation).
Nalu-Wind is a fully implicit scheme and requires a non-linear Picard iteration for each time
step, in this work we use two Picard iterations for every time step.
Each Picard iteration requires a Poisson solve and five Helmholtz solves
(three momentum equations, an enthalpy equation, and the subgrid-scale kinetic energy equation).
It is difficult to compare AMR-Wind and Nalu-Wind due to the differences in discretization and numerical
implementation, and additional study is required to understand the performance comparisons.
The most significant difference between the two solvers is that AMR-Wind uses structured grids
whereas Nalu-Wind uses unstructured grids. The differences in performance can be attributed with the
computational overhead of storing and manipulating unstructured data.

>>>>

Additional information about the mesh and domain requirements can be
drawn by examining the turbulent integral length scale $L$.  This
length scale can be calculated from $\langle R_{11}(\boldsymbol{\xi})
\rangle$ via
\begin{equation}
  L = \int_0^\infty \langle R_{11}(\xi) \rangle \: {\textrm d}\xi,
\end{equation}
and the results are shown in Table \ref{tab:GridStudyLscale}.  With a
resolution of 2.5 m, the ratio of the longitudinal length scale to the
grid size $L/\textrm{d}x \approx 10.1$, while in the lateral direction
$L/\textrm{d}x \approx 2.36$, suggesting even finer resolution is
required to capture the lateral scales.  This may be a critical
consideration when resolving both the lateral and longitudinal
turbulent scales are important, such as in SBL simulations with
turbine interactions.  However, in other situations, the additional
computational cost incurred may be prohibitive and only the
longitudinal component can be resolved.

@@

For both AMR-Wind and Nalu-Wind each time step is dominated by the linear solver work.
A time step within AMR-Wind requires two Poisson solves (MAC and nodal projection),
a tensor Helmholtz solve for Momentum equations, and two scalar Helmholtz solves
(potential temperature equation and subgrid-scale kinetic energy equation).
Nalu-Wind is a fully implicit scheme and requires a non-linear Picard iteration for each time
step, in this work we use two Picard iterations for every time step.
Each Picard iteration requires a Poisson solve and five Helmholtz solves
(three momentum equations, an enthalpy equation, and the subgrid-scale kinetic energy equation).
It is difficult to compare AMR-Wind and Nalu-Wind due to the differences in discretization and numerical
implementation, and additional study is required to understand the performance comparisons.
The most significant difference between the two solvers is that AMR-Wind uses structured grids
whereas Nalu-Wind uses unstructured grids. The differences in performance could be attributed with the
computational overhead of storing and manipulating unstructured data.
